### 2021-09-04 paper/amended.tex

\section{Amended Cryptol}\label{sec:amended}
Parker, will you start pulling your stuff in here