### 2021-09-04 paper/iota.tex

\section{Static Iota}\label{sec:iota}
Talk about changes we made to iota to prove out directly between precomputed and computed on the fly.