### 2021-09-11 paper/amended.tex
Update paper.tex and amended.tex Add Cryptol lstlisting highlighting.

\section{Amended Cryptol}
\label{sec:amended}
Initially it appeared that visual inspection with Cryptol’s standard syntax was subpar.
The language's functional folds and list comprehensions do not align easily with the imperative 
and array-based NIST standard. By leveraging Cryptol’s versatility, however, we produced a couple 
utility functions to visually mimic the sequential nature of the NIST specification.

\subsection{For Utility Function}
The phrase “for all” indicates a looping structure and is quite common in the NIST standard. 
It is used in the definition of the pi function shown in Figure \ref{fig:nistPi}.

\begin{figure}[h]
  \centering
  \begin{lstlisting}[basewidth = {.5em}]
1. For all triples (x,y,z) such that 
  0 <= x < 5, 0 <= y < 5, and 0 <= z < w,
  let A'[x, y, z] = A[(x + 3y) mod 5, x, z].
2. Return A'.
  \end{lstlisting}
  \caption{NIST Pi definition}
  \label{fig:nistPi}
\end{figure}

By nesting Cryptol's list comprehensions, we easily implemented the pi function in Figure 
\ref{fig:cryptolPi}.

\begin{figure}[h]
  \centering
\begin{lstlisting}[language=Cryptol]
type COL = 5 
type W = 64
type STATE_ARR = [COL][COL][W]
LIST4 = [0..COL - 1]
LIST63 = [0..W - 1]
pi : STATE_ARR -> STATE_ARR
pi a = [[[a @x @((x + 3*y) % 'COL) @z 
            | z <- LIST63] 
          | y <- LIST4] 
        | x <- LIST4]
\end{lstlisting}
\caption{Cryptol Pi definition}
\label{fig:cryptolPi}
\end{figure}

Cryptol declares list comprehension variables after they are used, and the order of the
nested loops is not as apparent. Using a helper function, ‘for’, we were able to provide
a more clear definition seen in Figure \ref{fig:cryptolAmendedPi}.

\begin{figure}[h]
  \centering
\begin{lstlisting}[language=Cryptol]
type COL = 5 
type W = 64
type STATE_ARR = [COL][COL][W]
LIST4 = [0..COL - 1]
LIST63 = [0..W - 1]
pi : STATE_ARR -> STATE_ARR
pi a = for LIST4 (&x ->
          for LIST4 (&y -> 
            for LIST4 (&z -> 
              a @x @((x + 3*y) % 'COL) @z)))
\end{lstlisting}
\caption{Amended Cryptol Pi definition}
\label{fig:cryptolAmendedPi}
\end{figure}

The ‘for’ utility function follows the same intuition as the implementation in most 
imperative languages in that it iterates and runs a loop body over defined elements. 
It is similar to the Array.prototype.map method in JavaScript in that it takes in a 
function and a list of arguments and returns a list of results from calling the function 
on every element of the input array. Figure \ref{fig:cryptolFor} shows it's implementation.

\begin{figure}[h]
  \centering
\begin{lstlisting}[language=Cryptol]
for : {n, a, b} [n]a -> (a -> b) -> [n]b
for vals loop = [loop index | index <- vals]
\end{lstlisting}
\caption{Cryptol For utility function}
\label{fig:cryptolFor}
\end{figure}

Though it is simple to implement, it reorders and structures the arguments in a very elegant way.
In constrast with standard cryptol syntax, the list to iterate over and the iterating 
variable are visible before the loop body. List comprehensions are hidden away behind this 
function call and visual equivalence is clearer.

\subsection{While Utility Function}
Another common phrasing is the “let...for...let”. This defines a state upon which a folding 
occurs, iteratively condensing the information into a final state. Imperatively, it is most 
comparable to a while loop. As shown by Figure \ref{fig:nistRC}, is present in the rc function 
defined by the NIST specification.

\begin{figure}[h]
  \centering
\begin{lstlisting}[basewidth = {.5em}]
1. If t mod 255 = 0, return 1.
2. Let R = 10000000.
3. For i from 1 to t mod 255, let:
  a. R = 0 || R;
  b. R[0] = R[0] ^ R[8];
  c. R[4] = R[4] ^ R[8];
  d. R[5] = R[5] ^ R[8];
  e. R[6] = R[6] ^ R[8];
  f. R =Trunc8[R].
4. Return R[0].
\end{lstlisting}
\caption{NIST RC definition}
\label{fig:nistRC}
\end{figure}

Maintaining this state through Cryptol's list comprehensions can be difficult.
The comprehensions require a list to iterate through.Since our ‘for’ utility 
function is simply a wrapper for list comprehensions, it is equally unsuited for 
visual verification on such steps. The condition to stop iterating in the RC function
bounds the number of iterations by [1, 254], which lets us construct such a list to 
iterate through. This leads to a rather ungainly Cryptol definition, however, as 
shown by Figure \ref{fig:cryptolRC}.

\begin{figure}[h]
  \centering
\begin{lstlisting}[language=Cryptol]
rc : [64] -> Bit
rc t = if (t % 255) == 0 
  then True 
  else rs !0 @0 where
    rs = [128:[8]] # [
      if i <= t % 255
        then (take'{8} ([
          (r@0) ^ (r@8),
          r@1,
          r@2,
          r@3, 
          (r@4) ^ (r@8),
          (r@5) ^ (r@8),
          (r@6) ^ (r@8),
          r@7,
          r@8])
            where r = [False] # (rs @(i - 1)))
        else rs @(i - 1)
      | i <- [1..254]]
\end{lstlisting}
\caption{Cryptol RC definition}
\label{fig:cryptolRC}
\end{figure}

Not only is this definition computationally slower as it runs 255 loops every time it is called, 
it also awkwardly indexes into a list of previous states. This differs vastly from the 
NIST specificiation both visually and functionally. Using a new ‘while’ utility function, we 
can declare an initial state, condition function, and body function that will recursively 
be called until the condition is met. It consolodates the information of the previous state 
until the condition is sastified. This definition is foudn in Figure \ref{fig:cryptolamendedRC}.

\begin{figure}[h]
  \centering
\begin{lstlisting}[language=Cryptol]
rc : [64] -> Bit
rc t = if (t % 255) == 0 
  then True 
  else (while {i = 1, R = 128:[8]}
    (&state -> state.i <= t % 255)
    (&state -> {
      i = state.i + 1, 
      R = (take'{8} ([
        (r@0) ^ (r@8),
        r@1,
        r@2,
        r@3,
        (r@4) ^ (r@8),
        (r@5) ^ (r@8),
        (r@6) ^ (r@8),
        r@7, 
        r@8]) 
          where r = [False] # state.R)})
    ).R @0
\end{lstlisting}
\caption{Amended Cryptol RC definition}
\label{fig:cryptolamendedRC}
\end{figure}

The final state is returned by the ‘while’ function. This iterates only until the 
condition is met, avoids the clutter of indexing within a larger array of all previous 
states, and places the condition in a more obvious location before the loop body. This 
utility function is defined in Figure \ref{fig:cryptolWhile}.

\begin{figure}[h]
  \centering
\begin{lstlisting}[language=Cryptol]
while : {a} a -> (a -> Bit) -> (a -> a) -> a
while state cond f = if (cond state)
  then (while (f state) cond f)	
  else state
\end{lstlisting}
\caption{Cryptol While utility function}
\label{fig:cryptolWhile}
\end{figure}