### 2021-09-16 paper/preliminaries.tex
Restructure sections

\section{Preliminaries}\label{sec:preliminaries}

\subsection{Tools}