### 2021-11-22 paper/paper.tex
Revise to use LLNCS formating for NFM Submission

\documentclass[runningheads]{llncs}
%
\usepackage{paralist}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{color}
\definecolor{blue}{RGB}{42,0,255}
\definecolor{green}{RGB}{63,127,95}
\definecolor{purple}{RGB}{127,0,85}
\definecolor{lime}{RGB}{114, 148, 0}
\usepackage{listings}
\lstdefinelanguage{Cryptol}{
  keywordstyle=[1]\color{blue}\bfseries,
  keywordstyle=[2]\color{purple}\bfseries,
  keywordstyle=[3]\color{black}\bfseries,
  basicstyle={\color{lime}\scriptsize},
  literate=*
  {*}{{{\color{black}$*$}}}{1}
  {=}{{{\color{black}=}}}{1}
  {+}{{{\color{black}+}}}{1}
  {-}{{{\color{black}$-$}}}{1}
  {\%}{{{\color{black}\%}}}{1}
  {(}{{{\color{black}(}}}{1}
  {)}{{{\color{black})}}}{1}
  {\{}{{{\color{black}\{}}}{1}
  {\}}{{{\color{black}\}}}}{1}
  {[}{{{\color{black}[}}}{1}
  {]}{{{\color{black}]}}}{1}
  {|}{{{\color{black}$|$}}}{1}
  {>}{{{\color{black}$>$}}}{1}
  {<}{{{\color{black}$<$}}}{1}
  {\#}{{{\color{black}\#}}}{1}
  {@}{{{\color{black}@}}}{1}
  {.}{{{\color{black}.}}}{1}
  {\'}{{{\color{black}'}}}{1}
  {:}{{{\color{black}:}}}{1}
  {^}{{{\color{black}$\char`\^$}}}{1}
  {&}{{{\color{black}\textbackslash}}}{1},
  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
  morekeywords=[2]{if, then, else, where},
  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
    vals, loop, state, while, take}
}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{almond}{rgb}{0.94, 0.87, 0.8}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}

\lstdefinestyle{customc}{
  frame=none,
  language=C,
  basicstyle=\footnotesize\ttfamily, 
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  numbers=left,
  keywordstyle=\bfseries\color{azure},
  commentstyle=\itshape\color{dkgreen},
  identifierstyle=\color{black},
  stringstyle=\color{almond},
  numbersep=-1pt,
  numberstyle=\color{gray},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
  Parker Hanson\inst{1} \and
  Benjamin Winters\inst{1} \and
  Eric Mercer\inst{1}\orcidID{0000--0002-2264-2958} \and
  Bret Decker\inst{1}
}
%
\authorrunning{P. Hanson et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
  Brigham Young University, Provo UT 84602, USA
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{abstract}
\keywords{verification \and cryptography \and symbolic execution \and SHA3.}
\end{abstract}

\section{Introduction}\label{sec:introduction}
\input{introduction}

%%\input{preliminaries}

\section{Proof Overview and Results}\label{sec:proof}
\input{proof}

\section{SHA3 FIPS Specification in Cryptol}\label{sec:fips}
\input{fips}

\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
\input{openssl}

\section{Results and Discussion}\label{sec:results}
\input{results}

\section{Related Work}\label{sec:related}
\input{related}

\section{Conclusion}\label{sec:conclusion}
\input{conclusion}

%
%
%
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{paper}
%
\end{document}