### 2021-11-22 paper/sha-3.tex
Restructure sections and file names

Overview the SHA-3 algorithm.
Add enough detail for the proof-outline and results to make sense in the next section.