### 2021-11-30 paper/paper.tex
Add background on SAW and SHA-3

\documentclass[runningheads]{llncs}
%
\usepackage{paralist}
\usepackage{comment}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\definecolor{blue}{RGB}{42,0,255}
\definecolor{green}{RGB}{63,127,95}
\definecolor{purple}{RGB}{127,0,85}
\definecolor{lime}{RGB}{114, 148, 0}
\usepackage{listings}
\lstdefinelanguage{Cryptol}{
  keywordstyle=[1]\color{blue}\bfseries,
  keywordstyle=[2]\color{purple}\bfseries,
  keywordstyle=[3]\color{black}\bfseries,
  basicstyle={\color{lime}\scriptsize},
  literate=*
  {*}{{{\color{black}$*$}}}{1}
  {=}{{{\color{black}=}}}{1}
  {+}{{{\color{black}+}}}{1}
  {-}{{{\color{black}$-$}}}{1}
  {\%}{{{\color{black}\%}}}{1}
  {(}{{{\color{black}(}}}{1}
  {)}{{{\color{black})}}}{1}
  {\{}{{{\color{black}\{}}}{1}
  {\}}{{{\color{black}\}}}}{1}
  {[}{{{\color{black}[}}}{1}
  {]}{{{\color{black}]}}}{1}
  {|}{{{\color{black}$|$}}}{1}
  {>}{{{\color{black}$>$}}}{1}
  {<}{{{\color{black}$<$}}}{1}
  {\#}{{{\color{black}\#}}}{1}
  {@}{{{\color{black}@}}}{1}
  {.}{{{\color{black}.}}}{1}
  {\'}{{{\color{black}'}}}{1}
  {:}{{{\color{black}:}}}{1}
  {^}{{{\color{black}$\char`\^$}}}{1}
  {&}{{{\color{black}\textbackslash}}}{1},
  morekeywords=[1]{Bit, Integer, STATE_ARR, W, COL, type, LIST4, LIST63, True, False},
  morekeywords=[2]{if, then, else, where},
  morekeywords=[3]{pi, rc, cond, a, b, x, y, z, for, n, r, R, f, rs, i, index, t, 
    vals, loop, state, while, take}
}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{almond}{rgb}{0.94, 0.87, 0.8}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}

\lstdefinestyle{customc}{
  frame=none,
  language=C,
  basicstyle=\footnotesize\ttfamily, 
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  numbers=left,
  keywordstyle=\bfseries\color{azure},
  commentstyle=\itshape\color{dkgreen},
  identifierstyle=\color{black},
  stringstyle=\color{almond},
  numbersep=-1pt,
  numberstyle=\color{gray},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand\keccak{\textsc{Keccak}}
\newcommand\shaThree{\texttt{SHA-3}}
\newcommand\nist{\texttt{NIST}}
\newcommand\fips{\texttt{FIPS 202}}
\newcommand\saw{\texttt{SAW}}
\newcommand\cryptol{Cryptol}
\newcommand\openssl{OpenSSL}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\algoref}[1]{Alg.~\ref{#1}}
\newcommand{\lineref}[1]{Line~\ref{#1}}

\begin{document}
%
\title{Verification of OpenSSL's SHA3 \keccak\ function using the Software Analysis Workbench}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
  Parker Hanson\inst{1} \and
  Benjamin Winters\inst{1} \and
  Eric Mercer\inst{1}\orcidID{0000-0002-2264-2958} \and
  Bret Decker\inst{1}
}
%
\authorrunning{P. Hanson et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
  Brigham Young University, Provo UT 84602, USA
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{abstract}
\keywords{verification \and cryptography \and symbolic execution \and SHA3.}
\end{abstract}

\section{Introduction}\label{sec:introduction}
\input{introduction}

\section{Background}\label{sec:background}
\cryptol\ and \saw\ are used to prove properties of \openssl's implementation of \shaThree\ relative to the \nist\ \fips\ publication \cite{fips202}. 
What the proofs actually say, and an understanding of their limitations, relies on (1) a basic knowledge of how \saw\ works and what \cryptol\ does relative to \saw, and (2) how the \shaThree\ algorithm constructs its digests.
Key elements relative to understanding the proof of each (\cryptol, \saw, and \shaThree) are defined in this section.

\subsection{Cryptol and SAW}\label{subsec:saw}
\input{saw}
\subsection{SHA-3 Overview}\label{subsec:sha3}
\input{sha-3}

\section{Proof Outline and Results}\label{sec:proof}
\input{proof}

\section{FIPS 202 Specification in Cryptol}\label{sec:fips}
\input{fips202}

\section{SHA3 OpenSSL in Cryptol}\label{sec:openssl}
\input{openssl}

\section{Related Work}\label{sec:related}
\input{related}

\section{Conclusion}\label{sec:conclusion}
\input{conclusion}

%
%
%
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{paper}
%
\end{document}