### 2022-01-06 paper/related.tex
Revises related.tex

The verification of OpenSSL's SHA2-256 implementation served as a direct predecessor to this work (!Citation2).
While SHA2 is also part of NIST's Secure Hash Algorithms, its internal structure differs vastly from the more complex and resistant SHA3.
The SHA2 proof, in turn, expanded the methodology by which Amazon's s2n HMAC was verified (!Citation2).
While SHA3 is not implemented in s2n, the HMAC verification utilized both a high-level and low-level cryptol specification in a similar manner to this work.
Both prior works compared equivalent C and Cryptol specification through SAW.